---
title: "Aides Drug"
author: "Anthony Falconio"
date: "put the date here"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

```{r include=FALSE}

drug_users <- filter(triUmp, DrugUser == "yes")

drug_users
```

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

The Dyplr tool called "filter" will be used to create a specific subset of the total data so that only the data where the actual drug users appear is showing rather than all of the people not included as a drug user. Then the CandP helper files will be used with R to display the following: numerical descriptive, graphical descriptive, and inferential results.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=drug_users)
barchartGC(~infected + treatment,data=drug_users, type="percent")

```
Two bar charts were created to compare the data between the people in the data where the first bar chart shows that the majority of the drug users who had aids were using a placebo, but the second chart shows the users of the actual drug have a lower percentage of aids when compared to the placebo which indicates that the drug is effective when compared to a placebo.
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=drug_users)
rowPerc(table1)
colPerc(table1)
```
In the data table, we can see the number of people in each group between those who have aids and those who don't who take the actual drug and those who take the placebo. By looking at the data, we can see that the number of aids cases in the placebo group are much higher than the number of cases in the group using the drug. In fact, the number of cases in the placebo group is among people infected with aids is 3 times higher than those who are taking the drug.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
```


```{r} 
chisqtestGC(table1)
```


```{r}  
fisher.test(table1)
```
Three seperate inferential tests were run in this section with varying levels of detail in the description of the data.
Firstly, the Chi-squared test mainly showed the p-value to highlight the significant difference in the data between placebo users and drug users and how the number of aids cases in the two groups reflected the effectiveness of the drug. The first test showed a p-value of 0.0000001242. The second test by Georgetown College showed a p-value of 0 and the Fisher exact test showed a p-value of 0.0000000423. All of the tests showed that their is a significant difference in the data in the number of aids cases between placebo users and drug users since all of the p-values were less than 0.05.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

